---
title: "Practical Machine Learning Course Project"
author: "PanMud"
date: "October 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preparation

Training data was obtained from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Testing data was obtained from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Libraries included for analysis

```{r, results='hide'}
library(ggplot2)
library(caret)
library(randomForest)
library(gbm)
library(doParallel)
library(dplyr)
library(e1071)
pmlTrain <- read.csv("pml-training.csv", head=TRUE, sep=",", na.strings=c("NA","#DIV/0!","")) 
pmlTest <- read.csv("pml-testing.csv", head=TRUE, sep=",", na.strings=c("NA","#DIV/0!",""))  
```

# Pre-processing and data splitting

```{r}

thresholdTrain <- sapply(pmlTrain, function(df) {sum(is.na(df)==TRUE)/length(df)})
thresholdTest <- sapply(pmlTest, function(df) {sum(is.na(df)==TRUE)/length(df)})

colidxTrain <-names(which(thresholdTrain<0.95))
pmlTrain<-pmlTrain[,colidxTrain]
colidxTest  <-names(which(thresholdTest<0.95))
pmlTest<-pmlTest[,colidxTest]

novTrain <- nearZeroVar(pmlTrain,saveMetrics=TRUE)
novTest <- nearZeroVar(pmlTest,saveMetrics=TRUE)
goodTrainData <- pmlTrain[,which(novTrain$nzv==FALSE)]
goodTestData <- pmlTest[,which(novTest$nzv==FALSE)]

rmInxTrain <- grepl("X|timestamp|user_name", names(goodTrainData))
goodTrainData <- goodTrainData[, which(rmInxTrain==FALSE)]

rmInxTest <- grepl("X|timestamp|user_name|problem_id", names(goodTestData))
goodTestData <- goodTestData[, which(rmInxTest==FALSE)]


set.seed(35161)
indexTrain <- createDataPartition (goodTrainData$classe, p=0.75, list=FALSE)
testing <-goodTrainData [- indexTrain,]
inTrain <- createDataPartition(testing$classe, p = 0.75)[[1]]
crossv_test <- testing[ -inTrain,]
training <- goodTrainData [indexTrain ,]
testing<-testing[inTrain,]

```

# Training the predictor using the Random Forest method

We use parallel processing to increase the speed
```{r}
 cl <- makeCluster(detectCores())
 registerDoParallel(cl)
 model1 <- train(classe ~ ., data=training, method="rf")
 pred1 <- predict(model1, testing)
 stopCluster(cl)
 plot(model1$finalModel)
```

## Confusion matrix

```{r}
 confusionMatrix(pred1, testing$classe)
```

The confusion matrix gives and accuracy of 99.69%
 
## Importance of predictors
 
 ```{r}
 print(plot(varImp(model1)))
 ```
## Out of sample error 
```{r}
pred1 <- predict(model1,crossv_test)
accuracy <- sum(pred1 == crossv_test$classe) / length(pred1)
accuracy
```
The out of sample Error achieved is  99.67 % with the validation set.
 
## Prediction of new values
 
The final step is to use the model and predict values from our test case
```{r}
 final<- predict(model1,goodTestData)
 final
```
 